---
title: "Targeting Based on Heterogeneous Treatment Effects - Step 1"
author: "Aziz Abdul Malik"
date: "March 2, 2018"
output:
  word_document: default
  pdf_document: default
graphics: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, eval = TRUE,
                      fig.width = 6, fig.height = 4.5, fig.align = "right", cache = TRUE)


```



```{r, eval=TRUE}
library(devtools)
install_github("walterwzhang/causalTree")
```

\bigskip


## Overview

Both the randomized treatment and the rich customer-level data are crucial to achieve the goals:

1. Predict **incremental dollar spending** that can be *attributed* to the catalog mailing. Formally, 
$$\tau_{i}=\mathbb{E}[Y_{i}(1)-Y_{i}(0)|\boldsymbol{x}_{i}],$$
the conditional average treatment effect of a catalog mailing. This heterogeneous treatment effect, $\tau_i$, is the expected change in spending due to the catalog mailing. In particular, I predict incremental dollar spending at the **customer-level**, based on all the features (variables) that capture past customer behavior.

2. Predict customer-level incremental profits due to targeting a customer (catalog mailing), and develop a targeting strategy.

3. Evaluate the predictive power and profitability of the targeting strategy. Compare different targeting strategies, based on different estimation methods.

4. Evaluate the predictive power and profitability of the targeting strategy that was developed using the 2015 data in an implementation of the strategy in 2016. This evaluation allows to assess the strict external validity or transportability of the results. 

\medskip


\newpage



## Step 1: Estimation and prediction of conditional average treatment effects

Using the 2015 data to estimate and validate several models to predict the heterogeneous treatment effects.

```{r}
library(bit64)
library(data.table)
library(glmnet)
library(causalTree)
library(ggplot2)
library(knitr)
```

```{r}
load("Customer-Development-2015.RData")
```

Split the sample into training and validation.

```{r, results = "hide"}
set.seed(2001)
crm_DT[, training_sample := rbinom(nrow(crm_DT), 1, 0.5)]

```


```{r}
setnames(crm_DT, "mailing_indicator", "W")
```


#### Data pre-processing

Remove highly correlated features from the data set


```{r}

cor_matrix = cor(crm_DT[, !c("customer_id", "W", "outcome_spend"),
with = FALSE])

cor_matrix[upper.tri(cor_matrix, diag = TRUE)] = NA

cor_DT = data.table(row = rep(rownames(cor_matrix), ncol(cor_matrix)),
                    col = rep(colnames(cor_matrix), each = ncol(cor_matrix)),
                    cor = as.vector(cor_matrix))
cor_DT = cor_DT[is.na(cor) == FALSE]

large_cor_DT = cor_DT[abs(cor) > 0.95]
kable(large_cor_DT, digits = 4)

crm_DT = crm_DT[, !large_cor_DT$row, with = FALSE]

```



#### Estimation of heterogeneous treatment effects

Using the training sample to estimate the conditional average treatment effect (CATE) on dollar spending, $\tau_i$, due to catalog targeting. 

Estimating **linear models with treatment-interactions**:

(a) OLS


(b) LASSO


Using a recently developed non-parametric estimator that directly predicts the CATE for each customer:

(c) Causal forest

\bigskip



#### Estimation

To simplify code I **created separate training and validation samples** from the full data set. 


```{r}
training_DT   = crm_DT[training_sample == 1,
                       !c("customer_id", "training_sample"), with = FALSE]
validation_DT = crm_DT[training_sample == 0,
                       !c("customer_id", "training_sample"), with = FALSE]
```

\bigskip



**a) OLS**


```{r, cache=TRUE}

fit_OLS = lm(outcome_spend ~ .*W , data = training_DT)
Summary_OLS<-summary(fit_OLS)

```

```{r, cache=TRUE}

DT_OLS_Predict_W_1_DT = copy(validation_DT)
DT_OLS_Predict_W_1_DT[, W:=1]

DT_OLS_Predict_W_0_DT <- copy(validation_DT)
DT_OLS_Predict_W_0_DT[, W:=0]

DT_OLS_Predict_W_1_DT[, pred_spend_W_1 := predict(fit_OLS, newdata = DT_OLS_Predict_W_1_DT)]
DT_OLS_Predict_W_0_DT[, pred_spend_W_0 := predict(fit_OLS, newdata = DT_OLS_Predict_W_0_DT)]

```



**b) LASSO**

```{r, cache=TRUE}

#LASSO - Train
X = model.matrix(outcome_spend ~ 0 + .*W, data = training_DT)
y = training_DT[, outcome_spend]

fit_LASSO = cv.glmnet(x = X, y = y, alpha = 1.0)

```



```{r, cache=TRUE}

#LASSO Predict
DT_LASSO_Predict_W_1_DT = copy(validation_DT)
DT_LASSO_Predict_W_1_DT[, W:=1]

DT_LASSO_Predict_W_0_DT <- copy(validation_DT)
DT_LASSO_Predict_W_0_DT[, W:=0]

X_new_LASSO_W1 = model.matrix(outcome_spend ~ 0 + .*W, data = DT_LASSO_Predict_W_1_DT)
X_new_LASSO_W0 = model.matrix(outcome_spend ~ 0 + .*W, data = DT_LASSO_Predict_W_0_DT)

DT_LASSO_Predict_W_1_DT[, pred_spend_W_1 := predict(fit_LASSO, newx = X_new_LASSO_W1, s = "lambda.min")]
DT_LASSO_Predict_W_0_DT[, pred_spend_W_0 := predict(fit_LASSO, newx = X_new_LASSO_W0, s = "lambda.min")]


```


**c) Causal Forest **

```{r, cache=TRUE}

set.seed(1964)
fit_Causal_Forest = causalForest(outcome_spend ~ .,
                   treatment = training_DT$W,
                   data = training_DT,
                   num.trees = 1000,
                   verbose = TRUE)

```


```{r, cache=TRUE}

#Causal Forest Predict
DT_Causal_Forest_Predict<-predict(fit_Causal_Forest, validation_DT)

```


#### Predict treatment effects



```{r, cache=TRUE}
predict_DT = validation_DT[, .(W, outcome_spend)]

#Conditional Average Treatment Effect --- OLS
predict_DT[, pred_T_E_OLS:= DT_OLS_Predict_W_1_DT$pred_spend_W_1 - DT_OLS_Predict_W_0_DT$pred_spend_W_0]
rm(DT_OLS_Predict_W_1_DT)
rm(DT_OLS_Predict_W_0_DT)

#Conditional Average Treatment Effect --- LASSO
predict_DT[, pred_T_E_LASSO:= DT_LASSO_Predict_W_1_DT$pred_spend_W_1 - DT_LASSO_Predict_W_0_DT$pred_spend_W_0]
rm(DT_LASSO_Predict_W_1_DT)
rm(DT_LASSO_Predict_W_0_DT)

predict_DT[, pred_T_E_Causal := DT_Causal_Forest_Predict]

head(predict_DT)

```



**Once all model predictions are added, saving predictions data to RData and model estimates in RDS files.**


```{r, cache=TRUE}

save(predict_DT, file = "predictions.RData")
saveRDS(fit_OLS, file = "fit_OLS")
saveRDS(fit_LASSO, file = "fit_LASSO")
saveRDS(fit_Causal_Forest, file = "fit_Causal_Forest")


```

